---
title: "Fundamentals of Programming October 2025"
author: "Berry Boessenkool"
date: now
date-format: YYYY-MM-DD, HH:mm
format:
  gfm:
    embed-resources: false
---

Hey everyone! Welcome to Fundamentals of Programming 2025/26!\
This is a github task in the course [FP25](https://open.hpi.de/courses/hpi-dh-fprog2025).

*note that it is fine to not really understand the code at this point - we'll get to that throughout the course :)*

Please go through the new [installation guide](https://github.com/brry/fpsetup#software-installation-guide) first!

## Get weather data

```{r, packages, warning=FALSE, message=FALSE}
if(!requireNamespace("rdwd", quietly=TRUE))
	install.packages("rdwd")
rdwd::updateRdwd()
```

download recent weather data using [rdwd](https://bookdown.org/brry/rdwd/)

```{r, get_weather_data}
#| message: false
library(rdwd)

# Define a local directory
locdir <- "~/DWDdata"

# Create it if it doesn't exist
if (!dir.exists(locdir)) {
  dir.create(locdir, recursive = TRUE)
}
link <- selectDWD("Potsdam", res="daily", var="kl", per="recent")

clim <- dataDWD(link,dir = locdir, varnames=TRUE, force=24)
```

## Visualise recent temperature

```{r, plot_clim}
plotDWD(clim, "TMK.Lufttemperatur", ylab = "mean air temperature (in °C)", col = "darkgreen")

```

## Visualise recent Wind Speed

```{r, plot_Wind_Speed}
plotDWD(clim, "FX.Windspitze")
```

## Transfer to Python

```{python, histograms, results="hide"}
clim_py = r.clim
import matplotlib
print(f"Dataset shape: {clim_py.shape[0]} rows, {clim_py.shape[1]} columns")
clim_py = clim_py.select_dtypes(include=['float64', 'int64'])
clim_py.hist(figsize=(20, 16), bins=5)
```

## Calculate Summary Statistics and Identify Extreme Days

```{python}
#| label: temperature-stats
#| echo: true

# Calculate and display some basic statistics
print("\n=== Summary Statistics for Temperature ===")
temp_stats = clim_py['TMK.Lufttemperatur'].describe()
print(temp_stats)

# Find and display the maximum and minimum temperature days
print("\n=== Extreme Temperature Days ===")
max_temp_idx = clim_py['TMK.Lufttemperatur'].idxmax()
min_temp_idx = clim_py['TMK.Lufttemperatur'].idxmin()

print(f"Hottest day: {r.clim.loc[max_temp_idx, 'MESS_DATUM']} with {clim_py.loc[max_temp_idx, 'TMK.Lufttemperatur']:.1f}°C")
print(f"Coldest day: {r.clim.loc[min_temp_idx, 'MESS_DATUM']} with {clim_py.loc[min_temp_idx, 'TMK.Lufttemperatur']:.1f}°C")
```

```{python, yazan_chunk, echo=FALSE}
print("Yazan added this chunk as part of the first homework")

# Uppercase letters
import string
alphabets = list(string.ascii_lowercase)

for i in range(65, 91):
    alphabets.append(chr(i))
print(alphabets)
print("wishing everyone a good start")
print("I love programming")
```
<<<<<<< HEAD

```{R, Mouna, Warning=FALSE}
# label: matrix-example
# Matrix
n = 1:25
n

mat = matrix(n, nrow = 5)
mat

```
=======
I am a beginner in programming and excited to learn R and Python this semester!



>>>>>>> 416bcaf463a102ed79886d065a59eac5591fdc32
